Estimating SARS-CoV-2 transmission in England from randomly sampled prevalence surveys

### 2024-03-27 biology-data.tex
finish proof reading
minor corrections / improvements

The time for each of these stages depends on various features of host and virus characteristics and has a large amount of variation.
In this thesis, I focus on pre-Alpha variants of SARS-CoV-2 in unvaccinated individuals.
The time from infection to infectiousness is known as the \emph{latent period}~\autocite{goughEstimation}.
The length of the latent period is hard to estimate because we rarely observe either the time of infection or the start of infectiousness.
Estimates of its length are based on careful contact tracing, where transmission times are thought to be known with a reasonable degree of certainty.
These estimates vary but suggest a mean length of 3--4 days for pre-Alpha variants~\autocite[and references therein]{zhaoEstimating}.
The time from infection to experiencing symptoms is known as the \emph{incubation period}; it is undefined for asymptomatic individuals~\autocite{portaEpiDictionary}.
As the symptom onset time (when symptoms start) is often known, this is easier to estimate than the latent period.
However, the unknown infection time still poses challenges, again requiring contact tracing data.
Therefore, estimates are sensitive to the choice of data and statistical model.
Its mean length is around 6 days~\autocite{wuIncubation,quesadaIncubation,aleneSerial}.

@@

The final step is poststratifying the incidence proportion to the CIS target population and subpopulations of interest.
To start, I draw a sample of 500 prevalence time series in each stratum from the approximate posterior distribution.
For each posterior draw and day, the incidence proportion in the (sub)population is the weighted average of the incidence proportion in each stratum, where the weights are the number of individuals in that stratum.
The population numbers for each stratum are provided by the Office for National Statistics.

>>>>

The time for each of these stages depends on various features of host and virus characteristics and has a large amount of variation.
In this thesis, I focus on pre-Alpha variants of SARS-CoV-2 in unvaccinated individuals.
The time from infection to infectiousness is known as the \emph{latent period}~\autocite{goughEstimation}.
The length of the latent period is hard to estimate because we rarely observe either the time of infection or the start of infectiousness.
Estimates of its length are based on careful contact tracing, where transmission times are thought to be known with a reasonable degree of certainty.
These estimates vary but suggest a mean length of 3--4 days for pre-Alpha variants~\autocite[and references therein]{zhaoEstimating}.
The time from infection to experiencing symptoms is known as the \emph{incubation period}; it is undefined for asymptomatic individuals~\autocite{portaEpiDictionary}.
As the symptom onset time (when symptoms start) is often known, this is easier to estimate than the latent period.
However, the unknown infection time still poses challenges, again requiring contact tracing data.
Therefore, estimates are sensitive to the choice of data and statistical model.
Its mean length is around 6 days~\autocite{wuIncubation,raiIncubation,aleneSerial}.

@@

The final step is poststratifying the incidence proportion to the CIS target population and subpopulations of interest.
To start, I draw a sample of 500 prevalence time series in each stratum from the approximate posterior distribution.
Denote these by $p^(j)_{t,i}$ for the $j$th sample where $i$ is a stratum, \ie some value of $a$, $e$, and $s$.
Define a (sub) population of interest as a set of strata $\set{S}$.
Then the prevalence in the (sub)population on day $t$ is
\begin{align}
p^{(j)}_t(\set{S})
&= \frac{\sum_{i \in \set{S}} p_{t,i} N_j}{\sum_{i \in \set{S}} N_i}
\end{align}
where $N_j$ is the population size of stratum $j$.
The population numbers for each stratum are provided by the Office for National Statistics.

### 2024-03-27 introduction.tex
minor corrections / improvements

Additionally, these models were used to explore policy scenarios.
For example, the decision at each of stage of lifting the lockdown in 2021 was informed by modelling studies~\autocite{sageEvidence}.
These aided decision-makers in understanding plausible impacts of their actions.

@@

The disease and observation models must be adapted to apply backcalculation to SARS-CoV-2 prevalence surveys (see \cref{E-inc-prev}).
The disease model must consider that this is now prevalence-type, rather than incidence-type data.
This difference induces a dependency structure between the observations on each day.
The observation model must now include the with error arising from using a sample of the population rather than (assumed) full recording of AIDS diagnoses.

>>>>

Additionally, these models were used to explore policy scenarios.
For example, the decision at each stage of lifting the lockdown in 2021 was informed by modelling studies~\autocite{sageEvidence}.
These aided decision-makers in understanding plausible impacts of their actions.

@@

The disease and observation models must be adapted to apply backcalculation to SARS-CoV-2 prevalence surveys (see \cref{E-inc-prev}).
The disease model must consider that this is now prevalence-type, rather than incidence-type data.
This difference induces a dependency structure between the observations on each day.
The observation model must now include the error arising from using a sample of the population rather than (assumed) full recording of AIDS diagnoses.

### 2024-03-27 transmission.tex
minor corrections / improvements

For each posterior sample of $\vec{Z}$, I poststratify it to the CIS target population and subpopulations of interest.
Specifically, define a (sub) population of interest as a set of strata $\set{S}$.
Then the incidence proportion in the subpopulation on day $t$ is
\begin{align}
Z_t(\set{S})
&= \frac{\sum_{j \in \set{S}} Z_{j,t} N_j}{\sum_{j \in \set{S}} N_j}
\label{backcalc:eq:poststratify}
\end{align}
where $N_j$ is the population size of stratum $j$.
The poststratified samples are now from the posterior of interest.

>>>>

For each posterior sample of $\vec{Z}$, I poststratify it to the CIS target population and subpopulations of interest, using the same basic method as in \cref{E-biology-data:sec:MRP}.
Specifically, define a (sub)population of interest as a set of strata $\set{S}$.
Then the incidence proportion in the (sub)population on day $t$ is
\begin{align}
Z_t(\set{S})
&= \frac{\sum_{j \in \set{S}} Z_{j,t} N_j}{\sum_{j \in \set{S}} N_j}
\label{backcalc:eq:poststratify}
\end{align}
where $N_j$ is the population size of stratum $j$.
The poststratified samples are now from the posterior of interest.
